---
title: "30DMC - Day 03 - Polygons"
format: html
editor: visual
---

## 30 Day Map Challenge 2024

Instructions: <https://30daymapchallenge.com/>

Day 03 is to make a map with **polygons** so I'm going to make maps of VT. I've done the world map, US map (which was terrible) and now will look for a VT map.

## Libraries and Prep

For this map, I'm going to map the biophysical regions of Vermont, and then overlay the biological hotspots. Source data for these come from:

-   <https://geodata.vermont.gov/datasets/VCGI::vt-biodiversity-project-biological-hotspots/about>

-   <https://geodata.vermont.gov/datasets/VTANR::vermont-biophysical-regions/about>

-   Do I need an outline of VT?

I wanted to create side by side choropleth maps to show time series data (2019 vs 2024 emerald ash borer sightings, for example) but I couldn't find the time series data, let alone by county or town. I'm sure it's out there, but I'm not a GIS specialist and this is not my full time job :)

::: callout-important
## This is the first map where I'm actually using shapefiles I downloaded. I also needed to install the ggspatial package based on Claude code. Also, the viridis package, which helps create color scales for color blind accessibility. Finally, used ggnewscale to make some Claude code work that (unsurprisingly) makes a new color scale for the legend.
:::

```{r}
#| message: false
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggrepel)
library(maps)
library(mapproj)
library(patchwork)
library(ggspatial)
library(viridis)
library(ggnewscale)

vt_outline <- st_read("C:/Users/ClintonSears/OneDrive - Bixal Solutions Inc/8 - R/30DMC_2024/Day 03 VT_State_Outline/FS_VCGI_OPENDATA_Boundary_BNDHASH_poly_vtbnd_SP_v1.shp")

biophysical <- st_read("C:/Users/ClintonSears/OneDrive - Bixal Solutions Inc/8 - R/30DMC_2024/Day 03 VT_Biophysical_Regions/Vermont_Biophysical_Regions.shp")

hotspots <- st_read("C:/Users/ClintonSears/OneDrive - Bixal Solutions Inc/8 - R/30DMC_2024/Day 03 VT_Biological_Hotspots/VT_Biodiversity_Project_-_Biological_Hotspots.shp")
```

::: {#tip-example .callout-tip}
## Hiding Unnecessary Code

The `echo: false` option disables the printing of code (only output is displayed). And message: false will suppress all (code and output)
:::

## Code from Claude

This was my prompt:

Can you help me generate some R code that will create a map of the USA. I want the map to outline states I have lived in in blue, which includes Vermont, New Hampshire, and Maine. And I want the map to outline states I have visited in green, which includes Florida, Illinois, and Michigan.

I had to fix depreciated code (use linewidth in place of size) and install additional packages, but it worked very well. Except I wanted just outline, not filled.

```{r}
# Create the map
ggplot() +
  # Base layer: Biophysical regions
  geom_sf(data = biophysical, 
          fill = "lightgreen", 
          color = "black") +
  
  # Add hotspots layer
  geom_sf(data = hotspots, 
          fill = "goldenrod", 
          alpha = 0.7) +
  
  # Add labels for biophysical regions
  geom_sf_text(data = biophysical,
               aes(label = NAME),  # Replace with your actual column name
               size = 3,
               check_overlap = TRUE) +
  
  # Add state outline
  geom_sf(data = vt_outline,
          fill = NA,
          color = "black",
          size = 1) +
  
  # Add scale bar and north arrow
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                        style = north_arrow_fancy_orienteering) +
  
  # Customize theme
  theme_minimal() +
  theme(
    legend.position = "left",
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.title = element_text(size = 10)
  ) +
  
  # Add title
  labs(title = "Vermont Biophysical Regions and Biological Hotspots",
       fill = "Features") +
  
  # Add legend
  scale_fill_manual(
    values = c("Biophysical Regions" = "lightgreen",
               "Biological Hotspots" = "goldenrod"),
    name = "Features",
    guide = guide_legend(title.position = "top")
  )

# Save the map if needed
# ggsave("vermont_map.png", width = 10, height = 12, units = "in", dpi = 300)
```

::: callout-caution
## No notes, but the map above needs additional work. Better labeled biophysical regions, no arrow, etc.
:::

## Modified Code from Claude

Prompt: Great! Thank you for that. Now please change the code so that the biophysical layer uses various colors for each polygon, and remove the north arrow. Also be sure that the legend shows the names of each biophysical region.

```{r}
# Create the map
ggplot() +
  # Base layer: Biophysical regions with distinct colors
  geom_sf(data = biophysical, 
          aes(fill = NAME),  # Replace with your actual column name
          color = "black") +
  
  # Add hotspots layer
  geom_sf(data = hotspots, 
          fill = "goldenrod", 
          alpha = 0.7) +
  
  # Add labels for biophysical regions
  geom_sf_text(data = biophysical,
               aes(label = NAME),  # Replace with your actual column name
               size = 3,
               check_overlap = TRUE) +
  
  # Add state outline
  geom_sf(data = vt_outline,
          fill = NA,
          color = "black",
          size = 1) +
  
  # Add scale bar
  annotation_scale(location = "br") +
  
  # Customize theme
  theme_minimal() +
  theme(
    legend.position = "left",
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.title = element_text(size = 10)
  ) +
  
  # Add title
  labs(title = "Vermont Biophysical Regions and Biological Hotspots",
       fill = "Biophysical Regions") +
  
  # Set colors for biophysical regions and add hotspots to legend
  scale_fill_viridis(
    discrete = TRUE,
    option = "turbo"  # This gives a nice range of distinct colors
  ) +
  # Add second legend for hotspots
  new_scale_fill() +
  geom_sf(data = hotspots,
          aes(fill = "Biological Hotspots"),
          alpha = 0.7) +
  scale_fill_manual(
    values = c("Biological Hotspots" = "goldenrod"),
    name = NULL
  )

# Save the map if needed
# ggsave("vermont_map.png", width = 10, height = 12, units = "in", dpi = 300)
```

## 

## My Final Code for Day 03

Changes include:

-   Center title

```{r}
# Create the map
ggplot() +
  # Base layer: Biophysical regions with distinct colors
  geom_sf(data = biophysical, 
          aes(fill = NAME),  # Replace with your actual column name
          color = "black") +
  
  # Add hotspots layer
  geom_sf(data = hotspots, 
          fill = "goldenrod", 
          alpha = 0.7) +
  
    # Add state outline
  geom_sf(data = vt_outline,
          fill = NA,
          color = "black",
          size = 3) +
  
  # Add scale bar
  annotation_scale(location = "br", height = unit(.5, "cm")) +
  
  # Customize theme
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0, size = 24, face = "bold"),
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 18),
    plot.subtitle = element_text(size = 12, face = "italic"),
    panel.grid = element_blank(),
    axis.text = element_blank(),
  ) +
  
  # Add title
  labs(x = NULL,
       y = NULL, 
       title = "Biological Hotspots within Vermont's Biophysical Regions",
       subtitle = "Hotspots note locations of \"rare, threatened, and endangered plants, animals, and exemplary natural communities\"",
       fill = "Biophysical Regions",
       caption = "Data accessed from Vermont Open Geodata Portal (https://geodata.vermont.gov/), 3 Nov 2024") +
  
  # Set colors for biophysical regions and add hotspots to legend
  scale_fill_viridis(
    discrete = TRUE,
    option = "turbo",  # This gives a nice range of distinct colors
    alpha = 0.5
  ) +
  # Add second legend for hotspots
  new_scale_fill() +
  geom_sf(data = hotspots,
          aes(fill = "Biological Hotspots"),
          alpha = 0.9) +
  scale_fill_manual(
    values = c("Biological Hotspots" = "gray10"),
    name = NULL
  )

# Save the map if needed
# ggsave("vermont_map.png", width = 10, height = 12, units = "in", dpi = 300)
```
