---
title: "30DMC - Day 02 - Lines"
format: html
editor: visual
---

## 30 Day Map Challenge 2024

Instructions: <https://30daymapchallenge.com/>

Day 02 is to make a map with **lines** which means roads, rivers, borders.

## Libraries and Prep

I'm building off of the same code I used for Day 01, and I've already got many packages installed.

For this map, I'm going to outline the US states that I have lived in, and maybe will outline in a second color those that I have visited.

::: callout-important
## I need to install a new package (maps) that will get US map data, as well as package (mapproj) for coordinates. Also, package (patchwork) for adding in Alaska and Hawaii.
:::

```{r}
#| message: false
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggrepel)
library(maps)
library(mapproj)
library(patchwork)

#points <- read.csv("C:/Users/ClintonSears/OneDrive - Bixal Solutions Inc/8 - R/30DMC_2024/Day 1 Points Data.csv", header = TRUE)
```

::: {#tip-example .callout-tip}
## Hiding Unnecessary Code

The `echo: false` option disables the printing of code (only output is displayed). And message: false will suppress all (code and output)
:::

## Code from Claude

This was my prompt:

Hello! Can you please help me generate R code to make a map of Vermont? I need to start with the outline of Vermont. Then I have two shapefiles I want to add to the map. The first shapefile shows the biophysical regions of Vermont. These need to be labeled with their names, and should all be colored light green. The second shapefile should go on top of the first and shows biological hotspots. These do not need labels and should be colored goldenrod. I'll need a legend on the left side of the map. Thanks!

```{r}
# Get US states map data
states_map <- map_data("state")

# Create vectors of states lived in and visited
lived_states <- c("vermont", "new hampshire", "maine")
visited_states <- c("florida", "illinois", "michigan")

# Create a data frame with state categories
state_categories <- data.frame(
  region = tolower(state.name),
  category = case_when(
    tolower(state.name) %in% lived_states ~ "Lived",
    tolower(state.name) %in% visited_states ~ "Visited",
    TRUE ~ "Other"
  )
)

# Merge the categories with the map data
states_map <- merge(states_map, state_categories, by = "region", all.x = TRUE)

# Create the map
ggplot(data = states_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = category), color = "black", linewidth = 0.2) +
  scale_fill_manual(
    values = c("Lived" = "lightblue", "Visited" = "lightgreen", "Other" = "white"),
    name = "State Status"
  ) +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ) +
  ggtitle("US States - Personal Travel History") +
  labs(caption = "Blue: States lived in | Green: States visited")
```

::: callout-caution
## What happened to AL, CA, CT, MS, etc.? There are weird polygons
:::

## My Code for 30DMC Day 02

Adapted from code above.

.

```{r}
# Get US states map data
states_map <- map_data("state")

# Create vectors of states lived in and visited
lived_states <- c("vermont", "new hampshire", "maine", "louisiana", "massachusetts", "georgia")
visited_states <- c("florida", "illinois", "michigan", "connecticut", "new york", "new jersey", "virginia", "north carolina", "indiana", "ohio", "pennsylvania", "california", "oregon", "washington", "wyoming", "colorado", "alabama", "mississippi", "maryland", "west virginia", "wisconsin", "tennessee")

# Create the categories for states
states_map$category <- case_when(
  states_map$region %in% lived_states ~ "Lived",
  states_map$region %in% visited_states ~ "Visited",
  TRUE ~ "Not Visited"
)

# Convert to factor with specific levels - order matters for drawing
states_map$category <- factor(states_map$category, 
                            levels = c("Not Visited", "Visited", "Lived"))

# Create the main map with layered drawing
main_map <- ggplot() +
  # Draw "Not Visited" states first
  geom_polygon(
    data = subset(states_map, category == "Not Visited"),
    aes(x = long, y = lat, group = group),
    fill = "white",
    color = "gray70",
    size = 1
  ) +
  # Draw "Visited" states second
  geom_polygon(
    data = subset(states_map, category == "Visited"),
    aes(x = long, y = lat, group = group),
    fill = "white",
    color = "gold",
    size = 1
  ) +
  # Draw "Lived" states last so they're on top
  geom_polygon(
    data = subset(states_map, category == "Lived"),
    aes(x = long, y = lat, group = group),
    fill = "white",
    color = "royalblue",
    size = 1
  ) +
  scale_color_manual(values = c("gray70", "gold", "royalblue"), 
                     labels = c("Not Visited", "Visited", "Lived")
                                                                                     ) +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  ) +
  ggtitle("US States - Lived, Visited, and Not Visited") +
  guides(color = guide_legend(title = "State Status"))
  
main_map
```

## 

## Suggested Improvement from Google Gemini

This works, but I am still missing two states + DC. Oh well.

```{r}
# Get US states map data
states_map <- map_data("state")

# Create vectors of states lived in and visited
lived_states <- c("vermont", "new hampshire", "maine", "louisiana", "massachusetts", "georgia")
visited_states <- c("florida", "illinois", "michigan", "connecticut", "new york", "new jersey", "virginia", "north carolina", "indiana", "ohio", "pennsylvania", "california", "oregon", "washington", "wyoming", "colorado", "alabama", "mississippi", "maryland", "west virginia", "wisconsin", "tennessee", "delaware")

# Create the categories for states
states_map$category <- case_when(
  states_map$region %in% lived_states ~ "Lived",
  states_map$region %in% visited_states ~ "Visited",
  TRUE ~ "Not Visited"
)

# Convert to factor with specific levels - order matters for drawing
states_map$category <- factor(states_map$category, 
                            levels = c("Not Visited", "Visited", "Lived"))

main_map <- ggplot() +
  geom_polygon(
    data = subset(states_map, category == "Not Visited"),
    aes(x = long, y = lat, group = group, color = "Not Visited"),
    fill = "white",
    size = 1
  ) +
  geom_polygon(
    data = subset(states_map, category == "Visited"),
    aes(x = long, y = lat, group = group, color = "Visited"),
    fill = "white",
    size = 1
  ) +
  geom_polygon(
    data = subset(states_map, category == "Lived"),
    aes(x = long, y = lat, group = group, color = "Lived"),
    fill = "white",
    size = 1
  ) +
  scale_color_manual(values = c("royalblue", "gray70", "gold"),
                     labels = c("Lived", "Not Visited", "Visited"),
                     guide = guide_legend(title = "Legend")) +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, face = "italic", color = "gray50"),
    legend.position = "bottom"
  ) +
  ggtitle("US States - Lived, Visited, and Not Visited")+
  labs(subtitle = "Where are Alaska, DC, and Hawaii??")

main_map
```
